Reinforcement Learning Produces Dominant Strategies for the Iterated Prisoner's Dilemma

### 2017-07-07 main.tex
Draw remaining diagrams for players.

The first strategy trained with reinforcement learning in the library is based
on lookup tables. The strategy encodes a set of deterministic responses
based on the opponent's first $n_1$ moves, the opponent's last $m_1$ moves, and
the players last $m_2$ moves. If $n_1 > 0$ then the player has infinite memory
depth, otherwise it has depth $\max{m_1, m_2}$. Although we tried various
combinations of $n_1, m_1,$ and $m_2$, the best performance at the time of
training was obtained for $n_1 = m_1 = m_2 = 2$ and generally for $n_1 > 0$.
The library includes a strategy
called EvolvedLookerUp2\_2\_2 which is among the top strategies in the library.

@@

\subsection{PSO Gambler}

PSO Gambler is a stochastic variant of LookerUp. Instead of deterministically
encoded moves the lookup table emits probabilities which are
used to choose cooperation or defection. The library includes a player trained
with $n_1 = m_1 = m_2 = 2$ that is \emph{mostly deterministic}, with most of the
probabilities being 0 or 1. At one time this strategy outperformed
EvolvedLookerUp2\_2\_2.

@@

Strategies based on artificial neural networks can also be trained with an
evolutionary algorithm. A variety of features are computed from the history
of play such as the opponents trailing moves, the total number of cooperations
of the player and the opponent, and several others, which are then input
into a feed forward neural network with one layer and user-supplied width.
An inner layer with just five nodes performs quite well in both deterministic and
noisy tournaments. The output of the ANN used in this work is deterministic;
a stochastic variant that outputs probabilities rather than exact moves could
be easily created.

@@

We used strategies based on finite state machines to create a number of
strategies. These strategies are deterministic and are efficient computationally.
In each round of play the strategy selects an action based on the current state
and the opponent's last action, transitioning to a new state for the next round.
These strategies can encode a variety
of other strategies, including classic strategies like TitForTat,
handshake strategies, and grudging strategies that always defect after
an opponent defection.

@@

We also trained stochastic versions of finite state machine players called
hidden Markov model players or HMMs. Like the finite state machines, these
strategies also encode an internal
state with probabilistic transitions based on the prior round of play to other
states and cooperate or defect with various probabilities at each state. These
are the best performing stochastic strategies in the library but take longer
to train due to their stochasticity.

>>>>

The first strategy trained with reinforcement learning in the library is based
on lookup tables. The strategy encodes a set of deterministic responses
based on the opponent's first $n_1$ moves, the opponent's last $m_1$ moves, and
the players last $m_2$ moves. If $n_1 > 0$ then the player has infinite memory
depth, otherwise it has depth $\max{m_1, m_2}$.
This is shown diagrammatically in Figure~\ref{fig:lookerup}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/lookerup}
    \caption{Diagrammatic representation of the Looker up Archetype}
    \label{fig:lookerup}
\end{figure}

Although we tried various
combinations of $n_1, m_1,$ and $m_2$, the best performance at the time of
training was obtained for $n_1 = m_1 = m_2 = 2$ and generally for $n_1 > 0$.
The library includes a strategy
called EvolvedLookerUp2\_2\_2 which is among the top strategies in the library.

@@

\subsection{Gambler}

Gambler is a stochastic variant of LookerUp. Instead of deterministically
encoded moves the lookup table emits probabilities which are
used to choose cooperation or defection.
This is shown diagrammatically in Figure~\ref{fig:gambler}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/gambler}
    \caption{Diagrammatic representation of the Gambler Archetype}
    \label{fig:gambler}
\end{figure}

The library includes a player trained
with $n_1 = m_1 = m_2 = 2$ that is \emph{mostly deterministic}, with most of the
probabilities being 0 or 1. At one time this strategy outperformed
EvolvedLookerUp2\_2\_2.

@@

Strategies based on artificial neural networks can also be trained with an
evolutionary algorithm. A variety of features are computed from the history
of play such as the opponents trailing moves, the total number of cooperations
of the player and the opponent, and several others, which are then input
into a feed forward neural network with one layer and user-supplied width.
This is shown diagrammatically in Figure~\ref{fig:ann}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/ann}
    \caption{Diagrammatic representation of the ANN Archetype}
    \label{fig:ann}
\end{figure}

An inner layer with just five nodes performs quite well in both deterministic and
noisy tournaments. The output of the ANN used in this work is deterministic;
a stochastic variant that outputs probabilities rather than exact moves could
be easily created.

@@

We used strategies based on finite state machines to create a number of
strategies. These strategies are deterministic and are efficient computationally.
In each round of play the strategy selects an action based on the current state
and the opponent's last action, transitioning to a new state for the next round.
These strategies can encode a variety
of other strategies, including classic strategies like TitForTat,
handshake strategies, and grudging strategies that always defect after
an opponent defection.
This is shown diagrammatically in Figure~\ref{fig:fsm}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/fsm}
    \caption{Diagrammatic representation of the Finite state machine Archetype}
    \label{fig:fsm}
\end{figure}

@@

We also trained stochastic versions of finite state machine players called
hidden Markov model players or HMMs. Like the finite state machines, these
strategies also encode an internal
state with probabilistic transitions based on the prior round of play to other
states and cooperate or defect with various probabilities at each state. These
are the best performing stochastic strategies in the library but take longer
to train due to their stochasticity.
This is shown diagrammatically in Figure~\ref{fig:hmm}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/hmm}
    \caption{Diagrammatic representation of the Hidden Markov Model Archetype}
    \label{fig:hmm}
\end{figure}
